\speculake: Hiding Malware in Speculative Execution

### 2018-05-07 paper/implementation.tex
add OpenSSL section

>>>>

Benign programs can act as a trigger program, provided they perform sufficient
indirect jumps to train the branch predictor. We experiment using the OpenSSL
library as a potential benign trigger, as its source code has gratuitous use of
function pointers which compile to indirect jumps. In addition, it has many
complicated code paths that can be easily selected by remote clients by their
choice of cipher suite.

For these experiments, we disable ASLR for simplification, but note that the
branch predictors can be used to determine ASLR offsets and our attack adjusted
accordingly~\cite{evtyushkin2016jump}.

We used \texttt{gdb} to run an instance of an OpenSSL server (version 1.0.1f),
and printed out every instruction executed and its address after a breakpoint on
the \texttt{SSL\_new} function. We then made a TLS connection to the
server\footnote{Which took well over 2~hours, potentially making it the slowest
TLS handshake to have ever completed},
which produced over 13~million instructions, including over 359,000 direct jumps
and 28,000 indirect jumps. We then searched for the longest repeated set of over
30 indirect jumps that end with a unique jump (i.e. source and destination do
not occur in the previous 30 indirect jumps).

We discovered a candidate that corresponds to code in OpenSSL's
\texttt{nistp256.c} that contained 31 indirect jumps repeated 254 times in our
handshake. This code is used during the TLS key exchange as the server computes
the ECDHE shared secret. We made a list of 31 source-destination address pairs
for these indirect jumps, and constructed a jump/ret chain at those addresses in
our payload program. Our program would jump to the source address of a pair,
which had a single \texttt{retq} instruction, which would pop the next
destination off the stack and perform the indirect jump. At each destination
address, we placed a static (direct) \texttt{jmpq} to bring execution to the
next source address in our list. The final jump source/destination pair was not
performed (and the destination was not pushed to the stack). However, we placed
code at the destination address that loaded a value into the cache (our target
function). The last pair's source performs an indirect jump that returns to our
payload program, which immediately times probing the cache to see if any value was
loaded.

We ran these experiments on an Intel Haswell
i5-4590 CPU, with OpenSSL and our payload program taskset to the same core.
We induced the jump pattern in OpenSSL by running Apache benchmark against it to
generate thousands of connections, using the ECDHE key exchange with the
secp256r1 curve (ECDHE-RSA-AES256-GCM-SHA384). When running Apache benchmark locally, our payload program
reliably achieves a 3.5\% hit rate. When apache benchmark runs on a remote
machine, the hit rate drops to approximately 2.0\%. Nonetheless, this is
sufficient to perform computation, as we simply perform thousands of iterations
to extract results from the speculative world before updating any real world state.

We verified that our payload program did not observe any cache loads when we ran
other programs that consumed CPU on the same core. In addition, when we used
Apache benchmark to create thousands of connections with a different cipher
suite (DHE-RSA-AES128-GCM-SHA256), we similarly received no hit rate. This
could potentially allow an adversary to use an obscure or uncommon (but still
supported) cipher suite to trigger a malicious payload program on a remote server.
